---
title: "Spectral Hypergraphs and Cell Fates - Code"
author: "Hugh Warden"
date: "14/04/2021"
output: 
    html_document:
      toc: true
      toc_float: true
      number_sections: true
      theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

This is a document explaining the computational methods used in "**insert paper title here**" by Raffaella Mulas, Hugh Warden, Michael Casey, Patrick Stumpf and Ben MacArthur.

All scripts and data sets can be found at https://github.com/hwarden162/SpectralHypergraphsAndCellFates

# Preprocessing

# MLModels

In this script the data is split for each individual, and for each individual is split into testing and trainiing data sets. Each of these training datasets is then used to train a multiple logistic regression classifier. Each of these classifiers is tested on each testing data set in turn and the softmax output and predicted classes are saved.

Firstly, the required libraries are loaded into R

```{r}
library(keras)
library(splitstackshape)
```

The keras library is used for machine learning and the splitstackshape library is for stratified sampling. Both of these will be covered in more detail in the relevant sections below.

Then the working directory is set, for this computer this is given by

```{r}
setwd("/Users/hugh/Documents/University/Maths/Year4/Project/RCode/MachineLearningCode/Demo")
```
## Preparing the Data

Then all of the preprocessed datasets from the previous step are loaded into the environment

```{r}
load("Data/MouseFeatures")
load("Data/HumanFeatures")
load("Data/MouseIdentity")
load("Data/HumanIdentity")
load("Data/MouseSpecimen")
load("Data/HumanSpecimen")
```

Mouse and human features are the processed feature sets from experiments and preprocessing of Stumpf et. Al. Mouse and human isentity contains the ground truth of those feature sets. Mouse and human specimen contains data scraped from the meta data of the feature sets of the Stumpf et Al experiment, containing an identifier as to which individual that observation came from.

These individual identifiers are then used to split the mouse and human features into smaller data sets for each individual.

```{r}
x.mouse <- list(spec1 = x.mouse[which(specimen.mouse == "M1"),], spec2 = x.mouse[which(specimen.mouse == "M2"),], spec3 = x.mouse[which(specimen.mouse == "M3"),], full = x.mouse)
x.human <- list(spec1 = x.human[which(specimen.human == "H1"),], spec2 = x.human[which(specimen.human == "H2"),], spec3 = x.human[which(specimen.human == "H3"),], full = x.human)
```

Then the data needs to be split into testing and training sets. For this experiment an 80/20 train/test split was used. To do this rows were randomly sampled using stratified sampling. This means that instead of randomly sampling 80% of the whole data set, the data set is split up by its class, and 80% of each class was selected to become part of the training data set. The indices are calculated via

```{r}
train_ind <- list(mouse = list(), human = list())
train_ind$mouse[[1]] <- stratified(data.frame(index = 1:nrow(x.mouse$spec1), identity = y.mouse[which(specimen.mouse == "M1")]), group = 2, size = 0.8)$index
train_ind$mouse[[2]] <- stratified(data.frame(index = 1:nrow(x.mouse$spec2), identity = y.mouse[which(specimen.mouse == "M2")]), group = 2, size = 0.8)$index
train_ind$mouse[[3]] <- stratified(data.frame(index = 1:nrow(x.mouse$spec3), identity = y.mouse[which(specimen.mouse == "M3")]), group = 2, size = 0.8)$index
train_ind$human[[1]] <- stratified(data.frame(index = 1:nrow(x.human$spec1), identity = y.human[which(specimen.human == "H1")]), group = 2, size = 0.8)$index
train_ind$human[[2]] <- stratified(data.frame(index = 1:nrow(x.human$spec2), identity = y.human[which(specimen.human == "H2")]), group = 2, size = 0.8)$index
train_ind$human[[3]] <- stratified(data.frame(index = 1:nrow(x.human$spec3), identity = y.human[which(specimen.human == "H3")]), group = 2, size = 0.8)$index
```

and then are used to further divide the feature data set

```{r}
x.mouse$spec1 <- list(train = x.mouse$spec1[train_ind$mouse[[1]],], test = x.mouse$spec1[-train_ind$mouse[[1]],])
x.mouse$spec2 <- list(train = x.mouse$spec2[train_ind$mouse[[2]],], test = x.mouse$spec2[-train_ind$mouse[[2]],])
x.mouse$spec3 <- list(train = x.mouse$spec3[train_ind$mouse[[3]],], test = x.mouse$spec3[-train_ind$mouse[[3]],])
x.human$spec1 <- list(train = x.human$spec1[train_ind$human[[1]],], test = x.human$spec1[-train_ind$human[[1]],])
x.human$spec2 <- list(train = x.human$spec2[train_ind$human[[2]],], test = x.human$spec2[-train_ind$human[[2]],])
x.human$spec3 <- list(train = x.human$spec3[train_ind$human[[3]],], test = x.human$spec3[-train_ind$human[[3]],])
```

The same principles are then used to split up the ground truth observations

```{r}
y.mouse.class <- list(spec1 = y.mouse[which(specimen.mouse == "M1")], spec2 = y.mouse[which(specimen.mouse == "M2")], spec3 = y.mouse[which(specimen.mouse == "M3")], full = y.mouse)
y.human.class <- list(spec1 = y.human[which(specimen.human == "H1")], spec2 = y.human[which(specimen.human == "H2")], spec3 = y.human[which(specimen.human == "H3")], full = y.human)

y.mouse.class$spec1 <- list(train = y.mouse.class$spec1[train_ind$mouse[[1]]], test = y.mouse.class$spec1[-train_ind$mouse[[1]]])
y.mouse.class$spec2 <- list(train = y.mouse.class$spec2[train_ind$mouse[[2]]], test = y.mouse.class$spec2[-train_ind$mouse[[2]]])
y.mouse.class$spec3 <- list(train = y.mouse.class$spec3[train_ind$mouse[[3]]], test = y.mouse.class$spec3[-train_ind$mouse[[3]]])
y.human.class$spec1 <- list(train = y.human.class$spec1[train_ind$human[[1]]], test = y.human.class$spec1[-train_ind$human[[1]]])
y.human.class$spec2 <- list(train = y.human.class$spec2[train_ind$human[[2]]], test = y.human.class$spec2[-train_ind$human[[2]]])
y.human.class$spec3 <- list(train = y.human.class$spec3[train_ind$human[[3]]], test = y.human.class$spec3[-train_ind$human[[3]]])
```

This process is then done again for the ground truth obserations, but first they are converted into one-hot encoding so that they can be used by the machine learning functions.

```{r}
y.mouse <- list(spec1 = to_categorical(as.numeric(y.mouse[which(specimen.mouse == "M1")])-1, 11), spec2 = to_categorical(as.numeric(y.mouse[which(specimen.mouse == "M2")])-1, 11), spec3 = to_categorical(as.numeric(y.mouse[which(specimen.mouse == "M3")])-1, 11), full = to_categorical(as.numeric(y.mouse)-1, 11))
y.human <- list(spec1 = to_categorical(as.numeric(y.human[which(specimen.human == "H1")])-1, 11), spec2 = to_categorical(as.numeric(y.human[which(specimen.human == "H2")])-1, 11), spec3 = to_categorical(as.numeric(y.human[which(specimen.human == "H3")])-1, 11), full = to_categorical(as.numeric(y.human)-1, 11))

y.mouse$spec1 <- list(train = y.mouse$spec1[train_ind$mouse[[1]],], test = y.mouse$spec1[-train_ind$mouse[[1]],])
y.mouse$spec2 <- list(train = y.mouse$spec2[train_ind$mouse[[2]],], test = y.mouse$spec2[-train_ind$mouse[[2]],])
y.mouse$spec3 <- list(train = y.mouse$spec3[train_ind$mouse[[3]],], test = y.mouse$spec3[-train_ind$mouse[[3]],])
y.human$spec1 <- list(train = y.human$spec1[train_ind$human[[1]],], test = y.human$spec1[-train_ind$human[[1]],])
y.human$spec2 <- list(train = y.human$spec2[train_ind$human[[2]],], test = y.human$spec2[-train_ind$human[[2]],])
y.human$spec3 <- list(train = y.human$spec3[train_ind$human[[3]],], test = y.human$spec3[-train_ind$human[[3]],])
```

The ground truth classes are then saved

```{r}
save(y.mouse.class, y.human.class, file = "Data/MLClasses.RData")
```

## Creating the MLR Models

A list is then made to store all of the machine learning models in

```{r}
ml.models <- list(mouse = list(), human = list())
```

The build_model function initialises and sets up a keras MLR model

```{r}
build_model <- function(in_shape) {
  model <- keras_model_sequential() %>%
    layer_dense(units = 11,
                activation = "softmax",
                input_shape = in_shape)
  model %>% compile(
    loss = "categorical_crossentropy",
    optimizer = "rmsprop",
    metrics = list("accuracy") )
  return(model)
}
```

This can then be used iteratively to generate an MLR model trained on each data set

```{r}
for (i in 1:3){
  ml.models$mouse[[i]] <- build_model(4311)
  ml.models$human[[i]] <- build_model(4311)
}
for (i in 1:3){
  ml.models$mouse[[i]] %>% fit(x.mouse[[i]]$train, y.mouse[[i]]$train)
  ml.models$human[[i]] %>% fit(x.human[[i]]$train, y.human[[i]]$train)
}
```

## Testing the MLR Models

Then each model is tested on the testing data. Two lists are generated to store these tests, one saves the softmax output of the classifier and one saves the predicted class.

```{r}
ml.sftmx <- list(mouse = list(list(), list(), list()), human = list(list(), list(), list()), mouse_human = list(list(), list(), list()), human_mouse = list(list(), list(), list()))
ml.class <- list(mouse = list(list(), list(), list()), human = list(list(), list(), list()), mouse_human = list(list(), list(), list()), human_mouse = list(list(), list(), list()))
for (i in 1:3){
  for (j in 1:3){
    ml.sftmx$mouse[[i]][[j]] <- ml.models$mouse[[i]] %>% predict_proba(x.mouse[[j]]$test)
    ml.sftmx$human[[i]][[j]] <- ml.models$human[[i]] %>% predict_proba(x.human[[j]]$test)
    ml.sftmx$mouse_human[[i]][[j]] <- ml.models$mouse[[i]] %>% predict_proba(x.human[[j]]$test)
    ml.sftmx$human_mouse[[i]][[j]] <- ml.models$human[[i]] %>% predict_proba(x.mouse[[j]]$test)
    
    ml.class$mouse[[i]][[j]] <- ml.models$mouse[[i]] %>% predict_classes(x.mouse[[j]]$test)
    ml.class$human[[i]][[j]] <- ml.models$human[[i]] %>% predict_classes(x.human[[j]]$test)
    ml.class$mouse_human[[i]][[j]] <- ml.models$mouse[[i]] %>% predict_classes(x.human[[j]]$test)
    ml.class$human_mouse[[i]][[j]] <- ml.models$human[[i]] %>% predict_classes(x.mouse[[j]]$test)
  }
}
```

Then all data and all of the models are saved

```{r}
save(ml.sftmx, ml.class, file = "Data/MLResults.RData")
save(x.mouse, y.mouse, x.human, y.human, train_ind, file = "Data/MLData.RData")
for (i in 1:3){
  save_model_tf(ml.models$mouse[[i]], file = paste0("Models/Mouse",i))
  save_model_tf(ml.models$human[[i]], file = paste0("Models/Human",i))
}
```

# SpecDists

This script takes the softmax outputs from the MLModels script and uses them to create hypergraphs. Then it finds the eigenvalues of the hypergraph normalised laplacian matrices and computes the spectral distance between all of the hypergraphs. It also computes the spectral distance from the theoretical disconnected hypergraph.

Firstly the required libraries are loaded

```{r}
library(RSpectra)
```

The RSpectra package is used as not all eigenvalues of the normalised laplacian matrix are calculated. For an mxn matrix only the first min(m,n) are needed as these are the only ones that can be non-zero and therefore effect the spectral distance.

Then the softmax data is loaded

```{r}
load("Data/MLResults.RData")
```

## Functions to Calculate Hypergraph Spectra

These functions are designed to calculate  the hyperedge normalised laplacian of a hypergraph given its incidence matrix, find its spectra and then compare these spectra.

### Hypergraph Degree Matrix

This is the function to find the hypergraph degree matrix

```{r}
get_degree_matrix <- function(inc_mat) {
  return(diag(nrow(inc_mat))*apply(inc_mat^2, 1, sum))
}
```

To do this it squares each element of the incidence matrix and then performs a row sum. Each of these elements is then put on the leading diagnal of an otherwise 0 matrix.

### Hyperedge Normalised Laplacian Matrix

This is the function to find the hyperedge normalised laplacian matrix of a hypergraph given its incidence matrix

```{r}
get_hyperedge_normalised_laplacian <- function(inc_mat) {
  deg_mat <- get_degree_matrix(inc_mat)
  diag(deg_mat) <- diag(deg_mat)^-1
  return(t(inc_mat)%*%deg_mat%*%inc_mat)
}
```

To do this it finds the degree matrix and then inverts it. This inversion should be turned into a pseudoinversion where zero entries remain as 0 so that isolated vertices don't create singular degree matrices. However, in this application it is not possible to have 0 entry as it is not in the image of the softmax function. In this case that means it is faster to use the true inverse as this can be vectorised easily.

### Eigenvalues of the Laplacian

This is the function to find the necessary eigenvalues of the laplacian matrix

```{r}
get_eigenvalues <- function(inc_mat) {
  inc_mat <- as.matrix(inc_mat)
  lap_mat <- get_hyperedge_normalised_laplacian(inc_mat)
  num <- min(dim(inc_mat))
  eigen <- eigs_sym(lap_mat, num)
  return(eigen$values)
}
```

The laplacian matrix for hypergraphs used has two forms, the vertex normalised laplacian and the hyperedge normalised laplacian. These both have the same multieigen set up to the multiplicity of the 0 eigenvalue, meaning that for an mxn incidence matrix they both have at most min(m,n) non-zero eigenvalues. As this analysis is using the spectral distance only non-zero eigenvalues matter with respect to analysis and therefore either normalised laplacian matrix can be calculated and then have the first min(m,n) eigenvalues taken.

### Spectral Distance

this is the function to calculate the spectral distance

```{r}
get_spectral_distance <- function(ev1, ev2) {
  ev1 <- sort(ev1, decreasing = TRUE)
  ev2 <- sort(ev2, decreasing = TRUE)
  n_1 <- length(ev1)
  n_2 <- length(ev2)
  if (n_1 > n_2) {
    diff <- n_1 - n_2
    ev2 <- c(ev2, rep(0, diff))
  }
  if (n_2 > n_1) {
    diff <- n_2 - n_1
    ev1 <- c(ev1, rep(0, diff))
  }
  ev <- sum(abs(ev1 - ev2))
  return(ev)
}
```

The spectral distance is calculated by summing up the absolute difference between respective eigenvalues. First the eigenvalues are therefore sorted into size order. Then a check is made that there are the same number of eigenvalues in each set and if not the smaller one is augmented with 0s. Then the spectral distance is calculated by summing absolute differences.

### Spectral Distance From Disconnected Hypergraph

This is the code to find the spectral distance from the disconnected hypergraph

```{r}
get_spectral_distance_disconnected <- function(ev){
  return(get_spectral_distance(ev, rep(1, length(ev))))
}
```

This is essentially a wrapper for the spectral distance function above that compares the spectra to that of the disconnected hypergraph which only has the non-zero eigenvalue 1, with arbitrary multiplicity.

## Creating Hypergraphs

Firstly, a list is created to store the hypergraphs so that they can be iterated over

```{r}
hypergraphs <- list(mouse_mouse = list(primal = list(list(), list(), list()), dual = list(list(), list(), list())), mouse_human = list(primal = list(list(), list(), list()), dual = list(list(), list(), list())), human_mouse = list(primal = list(list(), list(), list()), dual = list(list(), list(), list())), human_human = list(primal = list(list(), list(), list()), dual = list(list(), list(), list())))
```

To find the primal hyergraph (model 1) incidence matrix of the classifications made by training on Mouse i and testing on Human j would be found using 

```{r}
hypergraphs$mouse_human$primal[[i]][[j]]
```

These are then populated from the softmax outputs

```{r}
for (i in 1:3){
  for (j in 1:3){
    hypergraphs$mouse_mouse$primal[[i]][[j]] <- ml.sftmx$mouse[[i]][[j]]
    hypergraphs$mouse_human$primal[[i]][[j]] <- ml.sftmx$mouse_human[[i]][[j]]
    hypergraphs$human_mouse$primal[[i]][[j]] <- ml.sftmx$human_mouse[[i]][[j]]
    hypergraphs$human_human$primal[[i]][[j]] <- ml.sftmx$human[[i]][[j]]
    
    hypergraphs$mouse_mouse$dual[[i]][[j]] <- t(ml.sftmx$mouse[[i]][[j]])
    hypergraphs$mouse_human$dual[[i]][[j]] <- t(ml.sftmx$mouse_human[[i]][[j]])
    hypergraphs$human_mouse$dual[[i]][[j]] <- t(ml.sftmx$human_mouse[[i]][[j]])
    hypergraphs$human_human$dual[[i]][[j]] <- t(ml.sftmx$human[[i]][[j]])
  }
}
```

## Calculating Spectra

Again a list is created to store the spectra of each of the hypergraphs

```{r}
eigenvalues <- list(mouse_mouse = list(primal = list(list(), list(), list()), dual = list(list(), list(), list())), mouse_human = list(primal = list(list(), list(), list()), dual = list(list(), list(), list())), human_mouse = list(primal = list(list(), list(), list()), dual = list(list(), list(), list())), human_human = list(primal = list(list(), list(), list()), dual = list(list(), list(), list())))
```

The eigenvalues are then populated into the list using the functions defined above

```{r}
for (i in 1:3){
  for (j in 1:3){
    eigenvalues$mouse_mouse$primal[[i]][[j]] <- get_eigenvalues(hypergraphs$mouse_mouse$primal[[i]][[j]])
    eigenvalues$mouse_human$primal[[i]][[j]] <- get_eigenvalues(hypergraphs$mouse_human$primal[[i]][[j]])
    eigenvalues$human_mouse$primal[[i]][[j]] <- get_eigenvalues(hypergraphs$human_mouse$primal[[i]][[j]])
    eigenvalues$human_human$primal[[i]][[j]] <- get_eigenvalues(hypergraphs$human_human$primal[[i]][[j]])
    
    eigenvalues$mouse_mouse$dual[[i]][[j]] <- get_eigenvalues(hypergraphs$mouse_mouse$dual[[i]][[j]])
    eigenvalues$mouse_human$dual[[i]][[j]] <- get_eigenvalues(hypergraphs$mouse_human$dual[[i]][[j]])
    eigenvalues$human_mouse$dual[[i]][[j]] <- get_eigenvalues(hypergraphs$human_mouse$dual[[i]][[j]])
    eigenvalues$human_human$dual[[i]][[j]] <- get_eigenvalues(hypergraphs$human_human$dual[[i]][[j]])
  }
}
```

## Calculating Spectral Distances

A list is created to store all of the spectral distances

```{r}
spectral_distances <- list(primal = list(), dual = list(), disconnected = list())
```

### Primal and Dual Distances

The primal and dual spectral distances are then calculated and stored in the list

```{r}
spectral_distances$primal$mm <- list(mm = list(list(),list(),list()), mh = list(list(),list(),list()), hm = list(list(),list(),list()), hh = list(list(),list(),list()))
spectral_distances$primal$mh <- list(mm = list(list(),list(),list()), mh = list(list(),list(),list()), hm = list(list(),list(),list()), hh = list(list(),list(),list()))
spectral_distances$primal$hm <- list(mm = list(list(),list(),list()), mh = list(list(),list(),list()), hm = list(list(),list(),list()), hh = list(list(),list(),list()))
spectral_distances$primal$hh <- list(mm = list(list(),list(),list()), mh = list(list(),list(),list()), hm = list(list(),list(),list()), hh = list(list(),list(),list()))

spectral_distances$dual$mm <- list(mm = list(list(),list(),list()), mh = list(list(),list(),list()), hm = list(list(),list(),list()), hh = list(list(),list(),list()))
spectral_distances$dual$mh <- list(mm = list(list(),list(),list()), mh = list(list(),list(),list()), hm = list(list(),list(),list()), hh = list(list(),list(),list()))
spectral_distances$dual$hm <- list(mm = list(list(),list(),list()), mh = list(list(),list(),list()), hm = list(list(),list(),list()), hh = list(list(),list(),list()))
spectral_distances$dual$hh <- list(mm = list(list(),list(),list()), mh = list(list(),list(),list()), hm = list(list(),list(),list()), hh = list(list(),list(),list()))

spectral_distances$disconnected <- list(mm = list(), mh = list(), hm = list(), hh = list())
spectral_distances$disconnected$mm <- list(list(), list(), list())
spectral_distances$disconnected$mh <- list(list(), list(), list())
spectral_distances$disconnected$hm <- list(list(), list(), list())
spectral_distances$disconnected$hh <- list(list(), list(), list())

#Generating and storing all primal and dual spectral distances
for (i in 1:3){
  for (ii in 1:3){
    spectral_distances$primal$mm$mm[[i]][[ii]] <- list(list(),list(),list())
    spectral_distances$primal$mm$mh[[i]][[ii]] <- list(list(),list(),list())
    spectral_distances$primal$mm$hm[[i]][[ii]] <- list(list(),list(),list())
    spectral_distances$primal$mm$hh[[i]][[ii]] <- list(list(),list(),list())
    spectral_distances$dual$mm$mm[[i]][[ii]] <- list(list(),list(),list())
    spectral_distances$dual$mm$mh[[i]][[ii]] <- list(list(),list(),list())
    spectral_distances$dual$mm$hm[[i]][[ii]] <- list(list(),list(),list())
    spectral_distances$dual$mm$hh[[i]][[ii]] <- list(list(),list(),list())
    
    spectral_distances$primal$mh$mm[[i]][[ii]] <- list(list(),list(),list())
    spectral_distances$primal$mh$mh[[i]][[ii]] <- list(list(),list(),list())
    spectral_distances$primal$mh$hm[[i]][[ii]] <- list(list(),list(),list())
    spectral_distances$primal$mh$hh[[i]][[ii]] <- list(list(),list(),list())
    spectral_distances$dual$mh$mm[[i]][[ii]] <- list(list(),list(),list())
    spectral_distances$dual$mh$mh[[i]][[ii]] <- list(list(),list(),list())
    spectral_distances$dual$mh$hm[[i]][[ii]] <- list(list(),list(),list())
    spectral_distances$dual$mh$hh[[i]][[ii]] <- list(list(),list(),list())
    
    spectral_distances$primal$hm$mm[[i]][[ii]] <- list(list(),list(),list())
    spectral_distances$primal$hm$mh[[i]][[ii]] <- list(list(),list(),list())
    spectral_distances$primal$hm$hm[[i]][[ii]] <- list(list(),list(),list())
    spectral_distances$primal$hm$hh[[i]][[ii]] <- list(list(),list(),list())
    spectral_distances$dual$hm$mm[[i]][[ii]] <- list(list(),list(),list())
    spectral_distances$dual$hm$mh[[i]][[ii]] <- list(list(),list(),list())
    spectral_distances$dual$hm$hm[[i]][[ii]] <- list(list(),list(),list())
    spectral_distances$dual$hm$hh[[i]][[ii]] <- list(list(),list(),list())
    
    spectral_distances$primal$hh$mm[[i]][[ii]] <- list(list(),list(),list())
    spectral_distances$primal$hh$mh[[i]][[ii]] <- list(list(),list(),list())
    spectral_distances$primal$hh$hm[[i]][[ii]] <- list(list(),list(),list())
    spectral_distances$primal$hh$hh[[i]][[ii]] <- list(list(),list(),list())
    spectral_distances$dual$hh$mm[[i]][[ii]] <- list(list(),list(),list())
    spectral_distances$dual$hh$mh[[i]][[ii]] <- list(list(),list(),list())
    spectral_distances$dual$hh$hm[[i]][[ii]] <- list(list(),list(),list())
    spectral_distances$dual$hh$hh[[i]][[ii]] <- list(list(),list(),list())
    
    for (iii in 1:3){
      for (iv in 1:3){
        spectral_distances$primal$mm$mm[[i]][[ii]][[iii]][[iv]] <- get_spectral_distance(eigenvalues$mouse_mouse$primal[[i]][[ii]],eigenvalues$mouse_mouse$primal[[iii]][[iv]])
        spectral_distances$primal$mm$mh[[i]][[ii]][[iii]][[iv]] <- get_spectral_distance(eigenvalues$mouse_mouse$primal[[i]][[ii]],eigenvalues$mouse_human$primal[[iii]][[iv]])
        spectral_distances$primal$mm$hm[[i]][[ii]][[iii]][[iv]] <- get_spectral_distance(eigenvalues$mouse_mouse$primal[[i]][[ii]],eigenvalues$human_mouse$primal[[iii]][[iv]])
        spectral_distances$primal$mm$hh[[i]][[ii]][[iii]][[iv]] <- get_spectral_distance(eigenvalues$mouse_mouse$primal[[i]][[ii]],eigenvalues$human_human$primal[[iii]][[iv]])
        spectral_distances$dual$mm$mm[[i]][[ii]][[iii]][[iv]] <- get_spectral_distance(eigenvalues$mouse_mouse$dual[[i]][[ii]],eigenvalues$mouse_mouse$dual[[iii]][[iv]])
        spectral_distances$dual$mm$mh[[i]][[ii]][[iii]][[iv]] <- get_spectral_distance(eigenvalues$mouse_mouse$dual[[i]][[ii]],eigenvalues$mouse_human$dual[[iii]][[iv]])
        spectral_distances$dual$mm$hm[[i]][[ii]][[iii]][[iv]] <- get_spectral_distance(eigenvalues$mouse_mouse$dual[[i]][[ii]],eigenvalues$human_mouse$dual[[iii]][[iv]])
        spectral_distances$dual$mm$hh[[i]][[ii]][[iii]][[iv]] <- get_spectral_distance(eigenvalues$mouse_mouse$dual[[i]][[ii]],eigenvalues$human_human$dual[[iii]][[iv]])
        
        spectral_distances$primal$mh$mm[[i]][[ii]][[iii]][[iv]] <- get_spectral_distance(eigenvalues$mouse_human$primal[[i]][[ii]],eigenvalues$mouse_mouse$primal[[iii]][[iv]])
        spectral_distances$primal$mh$mh[[i]][[ii]][[iii]][[iv]] <- get_spectral_distance(eigenvalues$mouse_human$primal[[i]][[ii]],eigenvalues$mouse_human$primal[[iii]][[iv]])
        spectral_distances$primal$mh$hm[[i]][[ii]][[iii]][[iv]] <- get_spectral_distance(eigenvalues$mouse_human$primal[[i]][[ii]],eigenvalues$human_mouse$primal[[iii]][[iv]])
        spectral_distances$primal$mh$hh[[i]][[ii]][[iii]][[iv]] <- get_spectral_distance(eigenvalues$mouse_human$primal[[i]][[ii]],eigenvalues$human_human$primal[[iii]][[iv]])
        spectral_distances$dual$mh$mm[[i]][[ii]][[iii]][[iv]] <- get_spectral_distance(eigenvalues$mouse_human$dual[[i]][[ii]],eigenvalues$mouse_mouse$dual[[iii]][[iv]])
        spectral_distances$dual$mh$mh[[i]][[ii]][[iii]][[iv]] <- get_spectral_distance(eigenvalues$mouse_human$dual[[i]][[ii]],eigenvalues$mouse_human$dual[[iii]][[iv]])
        spectral_distances$dual$mh$hm[[i]][[ii]][[iii]][[iv]] <- get_spectral_distance(eigenvalues$mouse_human$dual[[i]][[ii]],eigenvalues$human_mouse$dual[[iii]][[iv]])
        spectral_distances$dual$mh$hh[[i]][[ii]][[iii]][[iv]] <- get_spectral_distance(eigenvalues$mouse_human$dual[[i]][[ii]],eigenvalues$human_human$dual[[iii]][[iv]])
        
        spectral_distances$primal$hm$mm[[i]][[ii]][[iii]][[iv]] <- get_spectral_distance(eigenvalues$human_mouse$primal[[i]][[ii]],eigenvalues$mouse_mouse$primal[[iii]][[iv]])
        spectral_distances$primal$hm$mh[[i]][[ii]][[iii]][[iv]] <- get_spectral_distance(eigenvalues$human_mouse$primal[[i]][[ii]],eigenvalues$mouse_human$primal[[iii]][[iv]])
        spectral_distances$primal$hm$hm[[i]][[ii]][[iii]][[iv]] <- get_spectral_distance(eigenvalues$human_mouse$primal[[i]][[ii]],eigenvalues$human_mouse$primal[[iii]][[iv]])
        spectral_distances$primal$hm$hh[[i]][[ii]][[iii]][[iv]] <- get_spectral_distance(eigenvalues$human_mouse$primal[[i]][[ii]],eigenvalues$human_human$primal[[iii]][[iv]])
        spectral_distances$dual$hm$mm[[i]][[ii]][[iii]][[iv]] <- get_spectral_distance(eigenvalues$human_mouse$dual[[i]][[ii]],eigenvalues$mouse_mouse$dual[[iii]][[iv]])
        spectral_distances$dual$hm$mh[[i]][[ii]][[iii]][[iv]] <- get_spectral_distance(eigenvalues$human_mouse$dual[[i]][[ii]],eigenvalues$mouse_human$dual[[iii]][[iv]])
        spectral_distances$dual$hm$hm[[i]][[ii]][[iii]][[iv]] <- get_spectral_distance(eigenvalues$human_mouse$dual[[i]][[ii]],eigenvalues$human_mouse$dual[[iii]][[iv]])
        spectral_distances$dual$hm$hh[[i]][[ii]][[iii]][[iv]] <- get_spectral_distance(eigenvalues$human_mouse$dual[[i]][[ii]],eigenvalues$human_human$dual[[iii]][[iv]])
        
        spectral_distances$primal$hh$mm[[i]][[ii]][[iii]][[iv]] <- get_spectral_distance(eigenvalues$human_human$primal[[i]][[ii]],eigenvalues$mouse_mouse$primal[[iii]][[iv]])
        spectral_distances$primal$hh$mh[[i]][[ii]][[iii]][[iv]] <- get_spectral_distance(eigenvalues$human_human$primal[[i]][[ii]],eigenvalues$mouse_human$primal[[iii]][[iv]])
        spectral_distances$primal$hh$hm[[i]][[ii]][[iii]][[iv]] <- get_spectral_distance(eigenvalues$human_human$primal[[i]][[ii]],eigenvalues$human_mouse$primal[[iii]][[iv]])
        spectral_distances$primal$hh$hh[[i]][[ii]][[iii]][[iv]] <- get_spectral_distance(eigenvalues$human_human$primal[[i]][[ii]],eigenvalues$human_human$primal[[iii]][[iv]])
        spectral_distances$dual$hh$mm[[i]][[ii]][[iii]][[iv]] <- get_spectral_distance(eigenvalues$human_human$dual[[i]][[ii]],eigenvalues$mouse_mouse$dual[[iii]][[iv]])
        spectral_distances$dual$hh$mh[[i]][[ii]][[iii]][[iv]] <- get_spectral_distance(eigenvalues$human_human$dual[[i]][[ii]],eigenvalues$mouse_human$dual[[iii]][[iv]])
        spectral_distances$dual$hh$hm[[i]][[ii]][[iii]][[iv]] <- get_spectral_distance(eigenvalues$human_human$dual[[i]][[ii]],eigenvalues$human_mouse$dual[[iii]][[iv]])
        spectral_distances$dual$hh$hh[[i]][[ii]][[iii]][[iv]] <- get_spectral_distance(eigenvalues$human_human$dual[[i]][[ii]],eigenvalues$human_human$dual[[iii]][[iv]])
      }
    }
  }
}
```

The spectral distance between the primal hypergraph trained on Mouse i and tested on Human j and the primal hypergraph trained on Human k and test on Human l can be found at

```{r}
spectral_distances$primal$mh$hh[[i]][[j]][[k]][[l]]
```

### Distances From the Disconnected Hypergraph

The spectral distances from the disconnected hypergraph were calculated by 

```{r}
for (i in 1:3){
  for (j in 1:3){
    spectral_distances$disconnected$mm[[i]][[j]] <- get_spectral_distance_disconnected(eigenvalues$mouse_mouse$dual[[i]][[j]])
    spectral_distances$disconnected$mh[[i]][[j]] <- get_spectral_distance_disconnected(eigenvalues$mouse_human$dual[[i]][[j]])
    spectral_distances$disconnected$hm[[i]][[j]] <- get_spectral_distance_disconnected(eigenvalues$human_mouse$dual[[i]][[j]])
    spectral_distances$disconnected$hh[[i]][[j]] <- get_spectral_distance_disconnected(eigenvalues$human_human$dual[[i]][[j]])
  }
}
```

To access the spectral distance from the disconnected hypergraph of the hypergraph trained on Human i and tested on Mouse j use the code

```{r}
spectral_distances$disconnected$hm[[i]][[j]]
```

# BalAcc

This script is used to calculate the confusion matrices and descriptive statistics (specifically balanced accuracies) of all of the classifiers.

First the libraries needed are loaded

```{r}
library(yardstick)
library(caret)
```

The yardstick library is used to calculate the balanced accuracy of classifications and the caret package is used to calculate the confusion matrix and various descriptive statistics of the classifications.

Than the data is loaded
```{r}
load("Data/MLResults.RData")
load("Data/MLClasses.RData")
```

Then lists are created to store the balanced accuracies and confidence matrices

```{r}
bal_accs <- list(mm = list(list(), list(), list()), mh = list(list(), list(), list()), hm = list(list(), list(), list()), hh = list(list(), list(), list()))

conf_mat <- list(mm = list(list(), list(), list()), mh = list(list(), list(), list()), hm = list(list(), list(), list()), hh = list(list(), list(), list()))
```

Then the confidence matrix and descriptive statistics are calculated for each set of classifications

```{r}
for (i in 1:3){
  for (j in 1:3){
    conf_mat$mm[[i]][[j]] <- confusionMatrix(as.factor(levels(y.mouse.class$full)[ml.class$mouse[[i]][[j]]+1]), y.mouse.class[[j]]$test)
    conf_mat$mh[[i]][[j]] <- confusionMatrix(as.factor(levels(y.mouse.class$full)[ml.class$mouse_human[[i]][[j]]+1]), y.human.class[[j]]$test)
    conf_mat$hm[[i]][[j]] <- confusionMatrix(as.factor(levels(y.human.class$full)[ml.class$human_mouse[[i]][[j]]+1]), y.mouse.class[[j]]$test)
    conf_mat$hh[[i]][[j]] <- confusionMatrix(as.factor(levels(y.human.class$full)[ml.class$human[[i]][[j]]+1]), y.human.class[[j]]$test)
  }
}
```

Then the balanced accuracy is calculated for each set of classifications

```{r}
for (i in 1:3){
  for (j in 1:3){
    bal_accs$mm[[i]][[j]] <- bal_accuracy(conf_mat$mm[[i]][[j]]$table)$.estimate
    bal_accs$mh[[i]][[j]] <- bal_accuracy(conf_mat$mh[[i]][[j]]$table)$.estimate
    bal_accs$hm[[i]][[j]] <- bal_accuracy(conf_mat$hm[[i]][[j]]$table)$.estimate
    bal_accs$hh[[i]][[j]] <- bal_accuracy(conf_mat$hh[[i]][[j]]$table)$.estimate
  }
}
```

# SDTables

This script is used to generate a matrix of spectral distances to compare as a heatmap.

Firstly two matrices are created to store the primal and spectral distances.

```{r}
sd.p <- matrix(0, 36,36)
sd.d <- matrix(0, 36,36)
```

Then the matrix is populated with the spectral distances

```{r}
sd.p[1:9,1:9] <- matrix(unlist(spectral_distances$primal$mm$mm), 9, 9, byrow = TRUE)
sd.d[1:9,1:9] <- matrix(unlist(spectral_distances$dual$mm$mm), 9, 9, byrow = TRUE)

sd.p[1:9,10:18] <- matrix(unlist(spectral_distances$primal$mm$mh), 9, 9, byrow = TRUE)
sd.d[1:9,10:18] <- matrix(unlist(spectral_distances$dual$mm$mh), 9, 9, byrow = TRUE)
sd.p[10:18,1:9] <- matrix(unlist(spectral_distances$primal$mm$mh), 9, 9, byrow = FALSE)
sd.d[10:18,1:9] <- matrix(unlist(spectral_distances$dual$mm$mh), 9, 9, byrow = FALSE)

sd.p[1:9,19:27] <- matrix(unlist(spectral_distances$primal$mm$hm), 9, 9, byrow = TRUE)
sd.d[1:9,19:27] <- matrix(unlist(spectral_distances$dual$mm$hm), 9, 9, byrow = TRUE)
sd.p[19:27,1:9] <- matrix(unlist(spectral_distances$primal$mm$hm), 9, 9, byrow = FALSE)
sd.d[19:27,1:9] <- matrix(unlist(spectral_distances$dual$mm$hm), 9, 9, byrow = FALSE)

sd.p[1:9,28:36] <- matrix(unlist(spectral_distances$primal$mm$hh), 9, 9, byrow = TRUE)
sd.d[1:9,28:36] <- matrix(unlist(spectral_distances$dual$mm$hh), 9, 9, byrow = TRUE)
sd.p[28:36,1:9] <- matrix(unlist(spectral_distances$primal$mm$hh), 9, 9, byrow = FALSE)
sd.d[28:36,1:9] <- matrix(unlist(spectral_distances$dual$mm$hh), 9, 9, byrow = FALSE)

sd.p[10:18,10:18] <- matrix(unlist(spectral_distances$primal$mh$mh), 9, 9, byrow = TRUE)
sd.d[10:18,10:18] <- matrix(unlist(spectral_distances$dual$mh$mh), 9, 9, byrow = TRUE)

sd.p[10:18,19:27] <- matrix(unlist(spectral_distances$primal$mh$hm), 9, 9, byrow = TRUE)
sd.d[10:18,19:27] <- matrix(unlist(spectral_distances$dual$mh$hm), 9, 9, byrow = TRUE)
sd.p[19:27,10:18] <- matrix(unlist(spectral_distances$primal$mh$hm), 9, 9, byrow = FALSE)
sd.d[19:27,10:18] <- matrix(unlist(spectral_distances$dual$mh$hm), 9, 9, byrow = FALSE)

sd.p[10:18,28:36] <- matrix(unlist(spectral_distances$primal$mh$hh), 9, 9, byrow = TRUE)
sd.d[10:18,28:36] <- matrix(unlist(spectral_distances$dual$mh$hh), 9, 9, byrow = TRUE)
sd.p[28:36,10:18] <- matrix(unlist(spectral_distances$primal$mh$hh), 9, 9, byrow = FALSE)
sd.d[28:36, 10:18] <- matrix(unlist(spectral_distances$dual$mh$hh), 9, 9, byrow = FALSE)

sd.p[19:27,19:27] <- matrix(unlist(spectral_distances$primal$hm$hm), 9, 9, byrow = TRUE)
sd.d[19:27,19:27] <- matrix(unlist(spectral_distances$dual$hm$hm), 9, 9, byrow = TRUE)

sd.p[19:27,28:36] <- matrix(unlist(spectral_distances$primal$hm$hh), 9, 9, byrow = TRUE)
sd.d[19:27,28:36] <- matrix(unlist(spectral_distances$dual$hm$hh), 9, 9, byrow = TRUE)
sd.p[28:36,19:27] <- matrix(unlist(spectral_distances$primal$hm$hh), 9, 9, byrow = FALSE)
sd.d[28:36,19:27] <- matrix(unlist(spectral_distances$dual$hm$hh), 9, 9, byrow = FALSE)

sd.p[28:36,28:36] <- matrix(unlist(spectral_distances$primal$hh$hh), 9, 9, byrow = TRUE)
sd.d[28:36,28:36] <- matrix(unlist(spectral_distances$dual$hh$hh), 9, 9, byrow = TRUE)
```

Then the row and column names are updated to represent the corresponding hypergraphs

```{r}
train_num <- rep(1:3, 4, each = 3)
test_num <- rep(1:3, 12)
train_spec <- rep(c("M", "H"), 1, each = 18)
test_spec <- rep(c("M", "H"), 2, each = 9)
id <- paste0(train_spec, train_num, "/", test_spec, test_num)

rownames(sd.p) <- id
rownames(sd.d) <- id
colnames(sd.p) <- id
colnames(sd.d) <- id
```

# Session Info

```{r, eval=TRUE}
sessionInfo()
```








